Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
library(dplyr)
pf <- read.delim('../data/pseudo_facebook.tsv')

qplot(x = age, y = friend_count, data = pf)

```

***

#### What are some things that you notice right away?
Response: People under 30 have more contacts on facebook than older people.

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(aes(x = age, y = friend_count), data = pf) +
    geom_point() + 
    xlim(13, 90)
```

***

### Overplotting
Notes: use alpha parameter and / or jitter to see numbers of points more clearly.

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) +
    geom_jitter(alpha = 1/20) + 
    xlim(13, 90)
```

#### What do you notice in the plot?
Response: The number of contacts dips in adulthood and begins to increase again slightly among middle aged adults. 

***

### Coord_trans()
Notes: transform the y coordinates to square root. note that you want to use point so that jitter does not jitter y to a negative value.

```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf) +
    geom_point(alpha = 1/20) + 
    xlim(13, 90) +
    coord_trans(y = "sqrt")
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(x = age, y = friend_count), data = pf) +
    geom_point(alpha = 1/20, position = position_jitter(h = 0)) + 
    xlim(13, 90) +
    coord_trans(y = "sqrt")
```

#### What do you notice?
Relatively few users over age 60 have fewer than 10 or so friends.

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(aes(x = age, y = friendships_initiated), data = pf) + 
    geom_jitter(alpha = 1/20) + 
    xlim(13, 90) 
    
```

***

### Overplotting and Domain Knowledge
Notes: Handy often to scale data to percentages when comparing groups with different absolute values of an activity of interest.

***

### Conditional Means
Notes: create a new frame with friend count summaries by age    

```{r Conditional Means}
pf.byAge <- pf %>%
    group_by(age) %>%
    summarize(mean_friend = mean(friend_count),
              median_friend = median(friend_count)) %>%
    arrange(age)

```

Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x = age, y = mean_friend), data = pf.byAge) + 
    geom_line() + 
    xlim(13, 90)
```

***

### Overlaying Summaries with Raw Data
Notes: layer summaries on the scatter

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friend_count), data = pf) +
    geom_point(alpha = 1/20, 
               position = position_jitter(h = 0),
               color = 'orange') + 
    geom_line(stat = 'summary', fun.y = mean) +
    geom_line(stat = 'summary', fun.y = quantile, fun.args= c(probs = 0.9),
              linetype = 2, color = 'blue') + 
    geom_line(stat = 'summary', fun.y = quantile, fun.args= c(probs = 0.1),
              linetype = 2, color = 'blue') +
    geom_line(stat = 'summary', fun.y = median,
              color = 'blue') +
    coord_trans(y = "sqrt", limx = c(13, 90)) 
```

#### What are some of your observations of the plot?
Response: the lower end of the contacts distribution increases with age -> older users are more likely to actually use their facebook account if they go through trouble of creating one.

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes: are age and friend count correlated?

```{r Correlation}
cor.test(pf$age, pf$friend_count, method = 'pearson')
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response: -0.0274

***

### Correlation on Subsets
Notes: what happens if you exclude respondents over "70"

```{r Correlation on Subsets}
with(subset(pf, age <= 70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes: Kendall and Spearman compare the rank orders of observations on two variables and provide a score. Spearman is the preferred method for measuring correlation between variables with ordinal values (likert scores, leaderboard rank, etc.). 

***

## Create Scatterplots
Notes: check out correlation between likes and web likes...

```{r}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + 
    geom_point(alpha = 1/20, color = 'orange') + 
    coord_cartesian(xlim=c(0, quantile(pf$www_likes_received, 0.95)),
                    ylim = c(0, quantile(pf$likes_received, 0.95))) + 
    geom_smooth(method = 'lm', color = 'blue')

```
***

### Strong Correlations
Notes: what is the pearson product moment coefficient?

```{r Strong Correlations}
with(pf, cor.test(pf$likes_received, pf$www_likes_received))
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
with(pf, cor.test(pf$likes_received, pf$www_likes_received))
```

Response: 0.948

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes: 

```{r More Caution With Correlation}
#install.packages('alr3')
library(alr3)
data("Mitchell")
```

Create your plot!

```{r Temp vs Month}
ggplot(aes(x = Temp, y = Month), data = Mitchell) + 
    geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
with(Mitchell, cor.test(x = Mitchell$Month, y = Mitchell$Temp))
```

***

### Making Sense of Data
Notes: organize the x axis into 12 month increments

```{r Making Sense of Data}
ggplot(aes(x = Month, y = Temp), data = Mitchell) + 
    geom_point() +
    scale_x_continuous(breaks = seq(0, 203, 12))
```

***

### A New Perspective

What do you notice?
Response: repeating pattern 

Watch the solution video and check out the Instructor Notes!
Notes: video suggests a t-test of independence of the variables as possible way to catch a non-linear relationship in the data. 
example: 
x <- seq(0, 4*pi, pi/20)
y <- cos(x)
qplot(x = x, y = y)
dcor.ttest(x, y)

does not work well with soil example (Mitchell)

google search shows another option is to look at distance correlation (aka Brownian covariance)... this does not work well with soil example either

dcov.test(x,y)

notes also suggest overlaying the years with nifty use of modulo:

ggplot(aes(x=(Month%%12),y=Temp),data=Mitchell)+
  geom_point()

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
pf$age_with_months <- (pf$age + ((12 - pf$dob_month) / 12))  
```

***

### Age with Months Means

```{r Age with Months Means}
pf.byMonth <- pf %>%
    group_by(age_with_months) %>%
    summarise(friend_count_mean = mean(friend_count),
              friend_count_median = median(friend_count),
              n = n()) %>%
    arrange(age_with_months)
```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(data = filter(pf.byAge, age < 71),
       aes(x = age, y = mean_friend)) +
    geom_line()

ggplot(data = filter(pf.byMonth, age_with_months < 71),
       aes(x = age_with_months, y = friend_count_mean)) +
    geom_line()
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
p1 <- ggplot(data = filter(pf.byAge, age < 71),
       aes(x = age, y = mean_friend)) +
    geom_line() +
    geom_smooth()

p2 <- ggplot(data = filter(pf.byMonth, age_with_months < 71),
       aes(x = age_with_months, y = friend_count_mean)) +
    geom_line() + 
    geom_smooth()

p3 <- ggplot(data = filter(pf, age < 71),
             aes(x = round(age / 5) * 5, y = friend_count))+
    geom_line(stat = "summary", fun.y = "mean")

library(gridExtra)
grid.arrange(p2, p1, p3, ncol = 1)
```

***

### Which Plot to Choose?
Notes: important to look at many plots in the discovery phase. when shifting to communication phase need to pick the plots that best tell the story.

***

### Analyzing Two Variables
Reflection: scatter plots are powerful tools for explorign relationships. the jitter and alpha features are both good tools for seeing patterns in a multitude of points, i.e. when data is over-plotted. Calculations based on variance such as correlation are useful measures for quantifying relationships between variables, unfortunately they are only effective at detecting linear relationships. Other measures of variance such as t-tests or distance correlation are better at finding non-linear relationships (though not good enough to find the pattern in soil temperature changes over time in the example.) Conditional summary statistics such as the mean help identify prevailing patterns in data. The choice of how data is grouped (e.g. by year, by month) influences how much noise remains in the data. Models that smooth out noise may not capture meaningful variations in data, so analysts are encouraged to explore the data many ways.  

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

