Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
library(ggplot2)
library(dplyr)
#bring in the sample dataset
pf <- read.delim('../data/pseudo_facebook.tsv')

#explore gender differences
ggplot(aes(x= gender, y= age),
       data= subset(pf, !is.na(gender))) + 
    geom_boxplot() + 
    stat_summary(fun.y= 'mean', geom = 'point', shape = 3) 

#by age
ggplot(aes(x= age, y= friend_count),
       data = subset(pf, !is.na(gender))) +
           geom_line(aes(color = gender), stat= 'summary', fun.y = median)

#create new frame with data grouped by age and gender
pf.fc_by_age_gender <- pf %>%
    filter(!is.na(gender)) %>%
    group_by(age, gender) %>%
    summarise(mean_friend_count = mean(friend_count),
              median_friend_count = median(friend_count),
              n = n()) %>%
    ungroup() %>%
    arrange(age)

```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
#plot median friend counts by gender by age
#recreate line plot from new frame
ggplot(aes(x= age, y= median_friend_count),
       data = pf.fc_by_age_gender) +
    geom_line(aes(color = gender))
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes: list median counts by gender in columns

```{r}
#install.packages('reshape2')
#library(reshape2)
#going to use tidyr instead
library(tidyr)

pf.fc_by_age_gender.wide <- pf.fc_by_age_gender %>%
    select(age, gender, median_friend_count) %>%
    spread(gender, median_friend_count)

```


***

### Ratio Plot
Notes: plot the ratio of female median friends to male median friends at each age

```{r Ratio Plot}
ggplot(data= pf.fc_by_age_gender.wide,
       aes(x=age, y= female / male)) +
    geom_line() +
    geom_hline(yintercept = 1, alpha = 0.3, linetype = 2)
```

***

### Third Quantitative Variable
Notes: going to add another variable to measure how long on facebook

```{r Third Quantitative Variable}
#first check dsitribution of tenure (figure out units)
qplot(data= pf, x= tenure)

pf$year_joined <- floor(2014 - (pf$tenure / 365))
summary(pf$year_joined)
table(pf$year_joined)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
pf$year_joined.bucket <- cut(pf$year_joined, breaks = c(2004, 2009, 2011, 2012, 2014))
```

***

### Plotting it All Together
Notes: one line for every year_joined bucket

```{r Plotting it All Together}
ggplot(data = subset(pf, !is.na(year_joined.bucket)),
       aes(x= age, y= friend_count)) +
    geom_line(aes(color= year_joined.bucket), 
              stat='summary', 
              fun.y = median)
```

***

### Plot the Grand Mean
Notes: change to mean and plot the overall mean

```{r Plot the Grand Mean}
ggplot(data = subset(pf, !is.na(year_joined.bucket)),
       aes(x= age, y= friend_count)) +
    geom_line(aes(color= year_joined.bucket), stat='summary', fun.y = mean) +
    geom_line(stat = 'summary', fun.y = mean, linetype = 2)
```

***

### Friending Rate
Notes: consider the rate at which people accumulate friends after joining. make sure they have at least one day tenure

```{r Friending Rate}
with(subset(pf, tenure >= 1), summary(friend_count / tenure))
```

***

### Friendships Initiated
Notes: plot the friendship rate for different tenure

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}
ggplot(data = subset(pf, tenure >= 1),
       aes(x= tenure, y= friendships_initiated / tenure)) +
    geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y=mean) +
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(data = subset(pf, tenure >= 1),
       aes(x= tenure, y= friendships_initiated / tenure)) +
    geom_smooth(aes(color = year_joined.bucket))

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}

```

***

### Number of Purchases
Notes:

```{r Number of Purchases}

```

***

### Prices over Time
Notes:

```{r Prices over Time}

```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}

```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

