---
title: "ProblemSet_GapMinder"
author: "Ryan Tuggle"
date: "March 29, 2016"
output: html_document
---

##Selecting Data Set
Reviewed files on GapMinder.Org
Focused on education category and selected spending on primary education.

```{r}
df <- read.csv("~/Repositories/data/expenditure_primary.csv")
summary(df)
```

Enriching dataset with region information from the country code library

```{r}
library(dplyr)
library(countrycode)
df <- countrycode_data %>% 
    select(country.name, continent, region) %>% 
    mutate(country = country.name) %>% 
    right_join(df, by = "country") %>%
    select(-country.name)

head(select(df, country, region, continent))                                                                                                                    
```

trying a different join method using regex for country
```{r}
for(r in c(1:length(df$country)){
    for(s in dt$gex)
    df$test[r] <-    
    flag <- 0
    while(flag == 0){
        
    }
}
```

Change shape of data to plot over time.

```{r}
library(tidyr)
prep <- gather(df, "year", "rate", X1998:X2011)
prep$year <- as.factor(gsub("X","",prep$year))
```

Box plots by region to see variation over time

```{r, echo=FALSE}
qplot(x = region, y = rate, data = subset(prep, !is.na(prep$continent)),
      geom = 'boxplot') +
    facet_wrap(~continent, scales = "free_x")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
