---
title: "problemset4"
author: "Ryan Tuggle"
date: "May 15, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem One: price vs. x

Plot price vs x in diamonds dataset.


```{r plot, echo = FALSE}
require(ggplot2)

ggplot(data = diamonds,
       aes(x = x, y = price)) +
    geom_point()
```

## Problem Two: Correlations

Find the correlation between price and x.

```{r correlation}
with(diamonds, cor.test(x = diamonds$x, y = diamonds$price))
cor.test(diamonds$y, y = diamonds$price)
cor.test(diamonds$z, diamonds$price)
```

##Problem Three: price vs. depth

Plot price vs. depth
make transparency 1/100 
explore x axis in units of two

```{r plot, echo = FALSE}

ggplot(data = diamonds,
       aes(x = depth, y = price)) + 
    geom_point(alpha = 1/100) +
    scale_x_continuous(limits = c(50,70), breaks = seq(50,70,2))
```

calculate the correlation

```{r correlation}
cor.test(diamonds$price, diamonds$depth)
```

##Problem Four: price vs. carat

plot price vs. carat
omit the top 1% of both

```{r plot}
require(dplyr)
ggplot(data = filter(diamonds, price <= quantile(price, probs = 0.99) & 
                         carat <= quantile(carat, probs = 0.99)),
       aes(x = carat, y = price)) +
    geom_point()

```


##Problem Five: Volume

add volume variable (x * y * z)
plot price and volume
calculate correlation excluding outliers at 0 and above 800

```{r volume}
diamonds$volume <- diamonds$x * diamonds$y * diamonds$z

ggplot(data = diamonds,
       aes(x = volume, y = price)) +
    geom_point()

with(filter(diamonds, volume < 800 & volume != 0), 
     cor.test(diamonds$volume, diamonds$price))

diamonds.sub <- filter(diamonds, volume < 800 & volume != 0)
cor.test(diamonds.sub$volume, diamonds.sub$price)

ggplot(data = subset(diamonds, volume < 800 & volume != 0),
       aes(x = volume, y = price)) +
    geom_point(alpha = 1/20) + 
    geom_smooth()

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
